\textit{Supplementary materials for}: Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-05-24 paper/main.tex

All of the statistical tests performed in our study were two-sided. The 95\%
confidence intervals we reported for each correlation were estimated from
bootstrap distributions of 10,000 correlation coefficients obtained by
sampling (with replacement) from the observed data.

@@

As illustrated in Figure~\ref{fig:sliding-windows}A, we start by building up a
corpus of documents using overlapping sliding windows that span each video's
transcript. Khan Academy provides professionally created, manual transcriptions
of all videos for closed captioning. However, such transcripts would not be
readily available in all contexts to which our framework could potentially be
applied. Khan Academy videos are hosted on the YouTube platform, which
additionally provides automated captions. We opted to use these automated
transcripts~\citep[which, in prior work, we have found to be of sufficiently
near-human quality to yield reliable data in behavioral
studies; ][]{ZimaEtal18} when developing our framework in order to make
it more directly extensible and adaptable by others in the future.

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. This $w$ parameter was chosen to match the same number of words
per sliding window (rounded to the nearest whole word, and before
preprocessing) as the sliding windows we defined in our prior
work~\citep[][i.e., 185 words per sliding window]{HeusEtal21}.

These sliding windows ramped up and down in length at the beginning and end of
each transcript, respectively. In other words, each transcript's first sliding
window covered only its first line, the second sliding window covered the first
two lines, and so on. This ensured that each line from the transcripts appeared
in the same number ($w$) of sliding windows. We next performed a series of
standard text preprocessing steps: normalizing case, lemmatizing, removing
punctuation and removing stop-words. We constructed our corpus of stop words by
augmenting the Natural Language Toolkit~\citep[NLTK; ][]{BirdEtal09} English
stop word list with the following additional words, selected using one of the
approaches suggested by~\citet{BoydEtal14}: ``actual,'' ``actually,'' ``also,''
``bit,'' ``could,'' ``e,'' ``even,'' ``first,'' ``follow,'' ``following,''
``four,'' ``let,'' ``like,'' ``mc,'' ``really,'', ``saw,'' ``see,'' ``seen,''
``thing,'' and ``two.'' This yielded sliding windows with an average of 73.8
remaining words, and lasting for an average of 62.22~seconds. We treated the
text from each sliding window as a single ``document'' and combined these
documents across the two videos' windows to create a single training corpus for
the topic model.

After fitting the topic model to the two videos' transcripts, we could use the
trained model to transform arbitrary (potentially new) documents into
$k$-dimensional topic vectors. A convenient property of these topic vectors is
that documents that reflect similar blends of topics (i.e., documents that
reflect similar themes, according to the model) will yield similar coordinates
(in terms of correlation, cosine similarity, Kullback-Leibler divergence,
Euclidean distance, or other geometric measures). In general, the similarity
between different documents' topic vectors may be used to characterize the
similarity in conceptual content between the documents.

We transformed each sliding window's text into a topic vector, and then used
linear interpolation (independently for each topic dimension) to resample the
resulting time series to one vector per second. We also used the fitted model to
obtain topic vectors for each question in our pool (see Supp.~Tab.~\questions).
Taken together, we obtained a \textit{trajectory} for each video, describing
its path through topic space, and a single coordinate for each question
(Fig.~\ref{fig:sliding-windows}C). Embedding both videos and all of the
questions using a common model enables us to compare the content from different
moments of videos, compare the content across videos, and estimate potential
associations between specific questions and specific moments of video.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints in the given lecture and all questions \textit{about} that
lecture appearing on the given quiz. We also define $f(s, \Omega)$ as the
$s$\textsuperscript{th} topic vector from the set of topic vectors $\Omega$.
Here $t$ indexes the set of lecture topic vectors $L$, and $i$ and $j$ index
the topic vectors of questions $Q$ used to estimate the knowledge trace. Note
that ``$\mathrm{correct}$'' denotes the set of indices of the questions the
participant answered correctly on the given quiz.

@@

In performing these analyses, our null hypothesis is that the knowledge
estimates we compute based on the quiz questions' embedding coordinates do
\textit{not} provide useful information about participants' abilities to answer
those questions---in other words, that there is no meaningful difference (on
average) between the knowledge estimates we compute for questions participants
answered correctly and those they answered incorrectly. Specifically, since we
estimate knowledge for a given embedding coordinate as a weighted
proportion-correct score (where each question's weight reflects its
embedding-space distance from the target coordinate; see Eqn.~\ref{eqn:prop}),
if these weights are uninformative (e.g., randomly distributed), then our
estimates of participants' knowledge should be equivalent (on average) to the
\textit{unweighted} proportion of correctly answered questions used to compute
them. In general, for a given participant and quiz, this expected value (i.e.,
that participant's proportion-correct score on that quiz) is the same for any
coordinate in the embedding space (e.g., any lecture timepoint, quiz question,
etc.). However, in the ``All questions'' and ``Within-lecture'' versions of the
analyses shown in Figure~\ref{fig:predictions}, we estimate each participant's
knowledge for each target question using all \textit{other} questions (or all
\textit{other} questions about the same lecture) they answered on the same quiz.
This introduces a systematic dependency between a participant's success on a
target question and their proportion-correct score on the remaining questions
available to estimate their knowledge for it. For example, suppose a participant
correctly answered $n$ out of $q$ questions on a given quiz. If we hold out a
single \textit{correctly} answered question as the target, the proportion of
remaining questions answered correctly would be $\frac{n - 1}{q - 1}$. Whereas
if we hold out a single \textit{incorrectly} answered question, the proportion
of remaining questions answered correctly would be $\frac{n}{q - 1}$. Thus, the
proportion of correctly answered remaining questions (and therefore the
null-hypothesized value of a knowledge estimate computed from them) is always
\textit{lower} for target questions a participant answered correctly than for
those they answered incorrectly.

To correct for this baseline inverse relationship between a participant's
success on a target question and their estimated knowledge for it, we used a
rebalancing procedure that ensured our knowledge estimates for questions each
participant answered correctly and incorrectly were computed from the
\textit{same} proportion of correctly answered questions. For each target
question on a given participant's quiz, we identified all remaining questions
with the opposite ``correctness'' label (i.e., if the target question was
answered correctly, we identified all remaining incorrectly answered questions,
and vice versa). We then held out each of these opposite-label questions, in
turn, along with the target question, and estimated the participant's knowledge
for the target question using all \textit{other} remaining questions. Since each
of these subsets of remaining questions was constructed by holding out one
correctly answered question and one incorrectly answered question from the
participant's quiz, if the participant correctly answered $n$ out of $q$
questions total, then their proportion-correct score on each subset of questions
used to estimate their knowledge for the target question is $\frac{n-1}{q-2}$,
regardless of whether they answered the target question correctly or
incorrectly. Finally, averaging over these per-subset knowledge estimates
yielded a rebalanced estimate of the participant's knowledge for the target
question that leveraged information from all remaining questions' embedding
coordinates, but whose expected value under our null hypothesis was the same as
that of each individual subset ($\frac{n-1}{q-2}$). By equalizing the
null-hypothesized values of knowledge estimates for correctly and incorrectly
answered questions, this procedure ensures that any meaningful relationships we
observe between participants' estimated knowledge for individual quiz questions
and their abilities to correctly answer them are attributable to the predictive
power of the embedding-space distances used to weight questions' contributions
to the knowledge estimates, rather than an artifact of our estimation procedure.
Note that if a participant answered all or no questions on a given quiz
correctly, their responses contained no opposite-label questions with which to
perform this rebalancing, and we therefore excluded their data from our analyses
for that quiz. We used this rebalancing procedure when constructing knowledge
estimates for the ``All questions'' and ``Within-lecture'' versions of the
analyses shown in Figure~\ref{fig:predictions}, but not for the
``Across-lecture'' analyses as, in this case, the target questions and the
questions used to estimate participants' knowledge for them were drawn from
different subsets of quiz questions (those about one lecture, and those about
the other), and were therefore independent.

@@

To assess the predictive value of our knowledge estimates, we compared each
GLMM's ability to explain participants' success on individual quiz questions to
that of an analogous model which assumed (as we assume under our null
hypothesis) that knowledge estimates for correctly and incorrectly answered
questions did \textit{not} systematically differ, on average. Specifically, we
used the same sets of observations with which we fit each ``full'' model to fit
a second ``null'' model that had the same random effects structure, but in which
the coefficient for the fixed effect of ''\texttt{knowledge}'' was fixed at 0
(i.e., we removed this term from the null model). We then compared each full
model to its reduced (null) equivalent using a likelihood-ratio test (LRT).
Because the standard asymptotic $\chi^2_d$ approximation of the null
distribution for the LRT statistic ($\lambda_{LR}$) can be anti-conservative for
finite sample sizes~\citep{GoldSimo00,ScheEtal08b,SnijBosk11}, we computed
$p$-values for these tests using a parametric bootstrap
procedure~\citep{DaviHink97,HaleHojs14}. For each of 10,000 bootstraps, we used
the fitted null model to simulate a sample of observations of equal size to our
original sample. We then re-fit both the null and full models to this simulated
sample and compared them via an LRT. This yielded a distribution of
$\lambda_{LR}$ statistics we may expect to observe given data that conforms to
our null hypothesis. We computed a corrected $p$-value for our observed
$\lambda_{LR}$ as $\frac{r+1}{n+1}$, where $r$ is the number of simulated model
comparisons that yielded a $\lambda_{LR}$ greater than our observed value and
$n$ is the number of simulations we ran (10,000).

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressible by the embedding
model---not solely the content explicitly probed by the quiz questions, or even
appearing in the lectures. To visualize these estimates
(Fig.~\ref{fig:knowledge-maps}, Supp.
Figs.~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used Uniform Manifold Approximation and
Projection~\citep[UMAP; ][]{McInEtal18a, McInEtal18b} to construct a 2D
projection of the text embedding space. Whereas our main analyses used a
15-topic embedding space, we used a 100-topic embedding space for these
visualizations. This change in the number of topics overcame an undesirable
behavior in the UMAP embedding procedure, whereby embedding coordinates for the
15-topic model tended to be ``clumped'' into separated clusters, rather than
forming a smooth trajectory through the 2D space. When we increased the number
of topics to 100, the embedding coordinates in the 2D space formed a smooth
trajectory through the space, with substantially less clumping
(Fig.~\ref{fig:knowledge-maps}). Creating a ``map'' by sampling this
100-dimensional space at high resolution to obtain an adequate set of topic
vectors spanning the embedding space would be computationally intractable.
However, sampling a 2D grid is trivial.

@@

To generate our estimates, we placed a set of 39 radial basis functions (RBFs)
throughout the embedding space, centered on the 2D projections for each
question (i.e., we included one RBF for each question). At coordinate $x$, the
value of an RBF centered on a question's coordinate $\mu$, is given by:
\begin{equation}
    \mathrm{RBF}(x, \mu, \lambda) = \exp\left\{-\frac{||x - \mu||^2}{\lambda}\right\}.
    \label{eqn:rbf}
\end{equation}
The $\lambda$ term in the RBF equation controls the ``smoothness'' of the
function, where larger values of $\lambda$ result in smoother maps. In our
implementation we used $\lambda = 50$.  Next, we estimated the ``knowledge''
at each coordinate, $x$, using:
\begin{equation}
    \hat{k}(x) = \frac{\sum_{i \in \mathrm{correct}} \mathrm{RBF}(x, q_i, \lambda)}{\sum_{j = 1}^N \mathrm{RBF}(x, q_j, \lambda)}.
    \label{eqn:rbf-knowledge}
\end{equation}
Equation~\ref{eqn:rbf-knowledge} computes the weighted proportion of correctly
answered questions, where the weights are given by how nearby (in the 2D space)
each question is to the $x$. We also defined \textit{learning maps} as the
coordinate-by-coordinate differences between any pair of knowledge maps.
Intuitively, learning maps reflect the \textit{change} in knowledge
across two maps.

>>>>

All of the statistical tests performed in our study were two-sided. The 95\%
confidence intervals we report for each correlation were estimated from
bootstrap distributions of 10,000 correlation coefficients obtained by
sampling (with replacement) from the observed data.

@@

As illustrated in Figure~\ref{fig:sliding-windows}A, we started by building up a
corpus of documents using overlapping sliding windows that spanned each lecture's
transcript. Khan Academy provides professionally created, manual transcriptions
of all lecture videos for closed captioning. However, such transcripts would not be
readily available in all contexts to which our framework could potentially be
applied. Khan Academy videos are hosted on the YouTube platform, which
additionally provides automated captions. We opted to use these automated
transcripts~\citep[which, in prior work, we have found to be of sufficiently
near-human quality to yield reliable data in behavioral
studies; ][]{ZimaEtal18} when developing our framework in order to make
it more directly extensible and adaptable by others in the future.

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. Each transcript
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the lecture (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. This $w$ parameter was chosen to match the same number of words
per sliding window (rounded to the nearest whole word, and before
preprocessing) as the sliding windows we defined in our prior
work~\citep[][i.e., 185 words per sliding window]{HeusEtal21}.

These sliding windows ramped up and down in length at the beginning and end of
each transcript, respectively. In other words, each transcript's first sliding
window covered only its first line, the second sliding window covered the first
two lines, and so on. This ensured that each line from the transcripts appeared
in the same number ($w$) of sliding windows. We next performed a series of
standard text preprocessing steps: normalizing case, lemmatizing, removing
punctuation and removing stop-words. We constructed our corpus of stop words by
augmenting the Natural Language Toolkit~\citep[NLTK; ][]{BirdEtal09} English
stop word list with the following additional words, selected using one of the
approaches suggested by~\citet{BoydEtal14}: ``actual,'' ``actually,'' ``also,''
``bit,'' ``could,'' ``e,'' ``even,'' ``first,'' ``follow,'' ``following,''
``four,'' ``let,'' ``like,'' ``mc,'' ``really,'', ``saw,'' ``see,'' ``seen,''
``thing,'' and ``two.'' This yielded sliding windows containing an average of 73.8
remaining words, and spanning an average of 62.22~seconds. We treated the
text from each sliding window as a single ``document'' and combined these
documents across the two lectures' windows to create a single training corpus for
the topic model.

After fitting the topic model to the two lectures' transcripts, we could use the
trained model to transform arbitrary (potentially new) documents into
$k$-dimensional topic vectors. A convenient property of these topic vectors is
that documents that reflect similar blends of topics (i.e., documents that
reflect similar themes, according to the model) will yield similar coordinates
(in terms of correlation, cosine similarity, Kullback-Leibler divergence,
Euclidean distance, or other geometric measures). In general, the similarity
between different documents' topic vectors may be used to characterize the
similarity in conceptual content between the documents.

We transformed each sliding window's text into a topic vector, and then used
linear interpolation (independently for each topic dimension) to resample the
resulting time series to one vector per second. We also used the fitted model to
obtain topic vectors for each quiz question in our pool (see Supp.~Tab.~\questions).
Taken together, we obtained a \textit{trajectory} for each lecture video, describing
its path through topic space, and a single coordinate for each question
(Fig.~\ref{fig:sliding-windows}C). Embedding both lectures and all of the
questions using a common model enables us to compare the content from different
moments of the lectures, compare the content across lectures, and estimate potential
associations between specific questions and specific moments of lecture content.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between the topic vectors for any lecture timepoint and quiz question, taken over all
timepoints in the given lecture and all questions \textit{about} that
lecture appearing on the given quiz. We also define $f(s, \Omega)$ as the
$s$\textsuperscript{th} topic vector from the set of topic vectors $\Omega$.
Here $t$ indexes the time series of lecture topic vectors $L$, and $i$ and $j$ index
the topic vectors of questions $Q$ used to estimate the participant's knowledge. Note
that ``$\mathrm{correct}$'' denotes the set of indices of the questions the
participant answered correctly on the given quiz.

@@

In performing these analyses, our null hypothesis is that the knowledge
estimates we compute based on the quiz questions' embedding coordinates do
\textit{not} provide useful information about participants' abilities to correctly answer
those questions---in other words, that there is no meaningful difference (on
average) between the knowledge estimates we compute for questions participants
answered correctly versus incorrectly. Specifically, since we
estimate knowledge for a given embedding coordinate as a weighted
proportion-correct score (where each question's weight reflects its
embedding-space distance from the target coordinate; see Eqn.~\ref{eqn:prop}),
if these weights are uninformative (e.g., randomly distributed), then our
estimates of participants' knowledge should be equivalent (on average) to the
\textit{unweighted} proportion of correctly answered questions used to compute
them. In general, for a given participant and quiz, this expected null value (i.e.,
that participant's proportion-correct score on that quiz) is the same for any
coordinate in the embedding space (e.g., any lecture timepoint, quiz question,
etc.). However, in the ``All questions'' and ``Within-lecture'' versions of the
analyses shown in Figure~\ref{fig:predictions}, we estimate each participant's
knowledge for each target question using all \textit{other} questions (or all
\textit{other} questions about the same lecture) they answered on the same quiz.
This introduces a systematic dependency between a participant's success on a
target question and their proportion-correct score on the remaining questions
available to estimate their knowledge for it. For example, suppose a participant
correctly answered $n$ out of $q$ questions on a given quiz. If we hold out a
single \textit{correctly} answered question as the target, the proportion of
remaining questions answered correctly would be $\frac{n - 1}{q - 1}$, whereas
if we hold out a single \textit{incorrectly} answered question, the proportion
of remaining questions answered correctly would be $\frac{n}{q - 1}$. Thus, the
proportion of correctly answered remaining questions (and therefore the
null-hypothesized value of a knowledge estimate computed from them) is always
\textit{lower} for target questions a participant answered correctly than for
those they answered incorrectly.

To correct for this baseline difference under our null hypothesis, we used a
rebalancing procedure that ensured our knowledge estimates for questions each
participant answered correctly and incorrectly were computed from the
\textit{same} proportion of correctly answered questions. For each target
question on a given participant's quiz, we first identified all remaining questions
with the opposite ``correctness'' label (i.e., if the target question was
answered correctly, we identified all remaining incorrectly answered questions,
and vice versa). We then held out each of these opposite-label questions, in
turn, along with the target question, and estimated the participant's knowledge
for the target question using all \textit{other} remaining questions. Since each
of these subsets of remaining questions was constructed by holding out one
correctly answered question and one incorrectly answered question from the
participant's quiz responses, if the participant correctly answered $n$ out of $q$
questions total, then their proportion-correct score on each subset of questions
used to estimate their knowledge would be $\frac{n-1}{q-2}$,
regardless of whether they answered the target question correctly or
incorrectly. Finally, we averaged over these per-subset knowledge estimates
to obtain a rebalanced estimate of the participant's knowledge for the target
question that leveraged information from all remaining questions' embedding
coordinates, but whose expected value under our null hypothesis was the same as
that of each individual subset ($\frac{n-1}{q-2}$). By equalizing the
null-hypothesized values of knowledge estimates for correctly and incorrectly
answered questions, this procedure ensures that any meaningful relationships we
observe between participants' estimated knowledge for individual quiz questions
and their abilities to correctly answer them reflect the predictive
power of the embedding-space distances we use to weight questions' contributions
to the knowledge estimates, rather than an artifact of our testing procedure.
Note that if a participant answered all or no questions on a given quiz
correctly, their responses contained no opposite-label questions with which to
perform this rebalancing, and we therefore excluded their data from our analyses
for that quiz. We used this rebalancing procedure when constructing knowledge
estimates for the ``All questions'' and ``Within-lecture'' versions of the
analyses shown in Figure~\ref{fig:predictions}, but not for the
``Across-lecture'' analyses as, in this case, the target questions and the
questions used to estimate participants' knowledge for them were drawn from
different subsets of quiz questions (those about one lecture, and those about
the other), and were therefore independent.

@@

To assess the predictive value of our knowledge estimates, we compared each
GLMM's ability to explain participants' success on individual quiz questions to
that of an analogous model which assumed (as we assume under our null
hypothesis) that knowledge estimates for correctly and incorrectly answered
questions did \textit{not} systematically differ, on average. Specifically, we
used the same sets of observations to which we fit each ``full'' model to fit
a second ``null'' model with the same random effects structure, but with
the coefficient for the fixed effect of ''\texttt{knowledge}'' constrained to zero
(i.e., we removed this term from the null model). We then compared each full
model to its reduced (null) equivalent using a likelihood-ratio test (LRT).
Because the standard asymptotic $\chi^2_d$ approximation of the null
distribution for the LRT statistic ($\lambda_{LR}$) can be anti-conservative for
finite sample sizes~\citep{GoldSimo00,ScheEtal08b,SnijBosk11}, we computed
$p$-values for these tests using a parametric bootstrap
procedure~\citep{DaviHink97,HaleHojs14}. For each of 10,000 bootstraps, we used
the fitted null model to simulate a sample of observations of equal size to our
original sample. We then re-fit both the null and full models to this simulated
sample and compared them via an LRT. This yielded a distribution of
$\lambda_{LR}$ statistics we may expect to observe given data that conforms to
our null hypothesis. We computed a corrected $p$-value for our observed
$\lambda_{LR}$ as $\frac{r+1}{n+1}$, where $r$ is the number of simulated model
comparisons that yielded a $\lambda_{LR}$ greater than our observed value and
$n$ is the number of simulations we ran (10,000).

@@

An important feature of our approach is that, given a trained text embedding
model and participants' performance on each quiz question, we can estimate
their knowledge about \textit{any} content expressible by the embedding
model---not solely the content explicitly probed by the quiz questions, or even
appearing in the lectures. To visualize these estimates
(Fig.~\ref{fig:knowledge-maps}, Supp.
Figs.~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used Uniform Manifold Approximation and
Projection~\citep[UMAP; ][]{McInEtal18a, McInEtal18b} to construct a 2D
projection of the text embedding space. Whereas our main analyses used a
15-topic embedding space, we used a 100-topic embedding space for these
visualizations. This change in the number of topics overcame an undesirable
behavior in the UMAP embedding procedure, whereby embedding coordinates for the
15-topic model tended to be ``clumped'' into separated clusters, rather than
forming a smooth trajectory through the 2D space. When we increased the number
of topics to 100, the embedding coordinates in the 2D space formed a smooth
trajectory through the space, with substantially less clumping
(Fig.~\ref{fig:knowledge-maps}). Creating a ``map'' by sampling this
100-dimensional space at high resolution to obtain an adequate set of topic
vectors spanning the embedding space would be computationally intractable.
However, sampling a 2D grid is trivial.

@@

To generate our estimates, we placed a set of 39 radial basis functions (RBFs)
throughout the embedding space, centered on the 2D projections for each
question (i.e., we included one RBF for each question). At coordinate $x$, the
value of an RBF centered on a question's coordinate $\mu$ is given by:
\begin{equation}
    \mathrm{RBF}(x, \mu, \lambda) = \exp\left\{-\frac{||x - \mu||^2}{\lambda}\right\}.
    \label{eqn:rbf}
\end{equation}
The $\lambda$ term in the RBF equation controls the ``smoothness'' of the
function, where larger values of $\lambda$ result in smoother maps. In our
implementation we used $\lambda = 50$.  Next, we estimated the ``knowledge''
at each coordinate, $x$, using:
\begin{equation}
    \hat{k}(x) = \frac{\sum_{i \in \mathrm{correct}} \mathrm{RBF}(x, q_i, \lambda)}{\sum_{j = 1}^N \mathrm{RBF}(x, q_j, \lambda)}.
    \label{eqn:rbf-knowledge}
\end{equation}
Equation~\ref{eqn:rbf-knowledge} computes the weighted proportion of correctly
answered questions, where the weights are given by how nearby (in the 2D space)
each question is to the $x$. We also defined \textit{learning maps} as the
coordinate-by-coordinate differences between any pair of knowledge maps.
Intuitively, learning maps reflect the \textit{change} in knowledge
across two maps.

### 2025-05-24 paper/supplement.tex

We also examined both models' abilities to match the quiz questions to specific 
temporal intervals within the lectures. The third and fourth rows of 
Supplementary Figure~\ref{fig:compare-bert} display the time series of 
correlations between the embedding weights for each quiz question and each 
timepoint of its corresponding lecture (third row: \textit{Four Fundamental 
Forces}; fourth row: \textit{Birth of Stars}; analogous to main text 
Fig.~\questionCorrs). When computed using LDA embeddings, the questions' 
correlation time series generally appear fairly different from each other (i.e., 
individual questions' traces can be easily visually separated), whereas their 
correlation time series computed using BERT embeddings are all highly similar. 
This indicates that the LDA embeddings are capturing some differences in content 
between the questions about a given lecture that the BERT embeddings are not. 
The time series plots of individual questions' correlations also exhibit clear 
``peaks'' when computed using LDA embeddings, but not when computed using BERT 
embeddings. This indicates that the LDA embeddings are capturing specific 
``matches'' between individual quiz questions and relatively well defined time 
periods of lecture content, whereas the BERT embeddings appear to blur these 
specific correspondences. Taken together, these comparisons suggest that our LDA 
model trained specifically on the two lectures' contents is sensitive to subtle 
differences between the concepts probed by individual quiz questions, and can 
identify the relevant periods in the lectures where those concepts where 
presented (Fig.~\questionCorrs, 
Supp.~Figs.~\ref{fig:forces-peaks},~\ref{fig:bos-peaks}, 
Supp.~Tab.~\ref{tab:matches}). Meanwhile, BERT, whose embedding weights must be 
able to represent a far broader range of concepts, consequently blurs these 
subtle distinctions between concepts conveyed within the span of a single course 
lecture, thereby affording a ``lower resolution'' depiction of its contents.

@@

These patterns are reflected in the similarity time series plots shown in panels 
C and D (analogous to main text Fig.~\questionCorrs). The questions' time series 
of topic vector correlations exhibit visually distinct ``peaks'' during specific 
sections of the lectures, outside of which their correlations tend to be 
relatively low. In other words, the set of timepoints whose topic vectors are 
maximally correlated with the topic vector for a given question tend to be not 
only specific, but also temporally contiguous. This indicates that topic vector 
correlation matches questions with sustained, cohesive periods of lecture 
content where the relevant underlying \textit{concepts} are discussed, rather 
than with a sporadic assortment of timepoints where the question and lecture 
content share more superficial similarities. The time series of Jaccard 
similarities, by contrast, are less cleanly structured and do appear to reflect 
such superficial content matches in the more scattered, brief, and weak 
``peaks'' they exhibit. As an example, panel E displays text of the two quiz 
questions highlighted in green and purple in panels A through D, along with the 
text from the best-matching intervals from the lectures (identified based on
topic vector correlation). Both questions include words that---given the subject 
matter of the two lectures---appear frequently throughout their transcripts 
(e.g., ``\textit{force}'', ``\textit{atom}'', ``\textit{nucleus}'', 
``\textit{cloud}'', ``\textit{dense}'', etc.). While the presence or absence of 
any one (or more) of these words will lead to differences in the degree of 
``match'' based on Jaccard similarity (resulting in the ``jagged'' time series
shown in panel D), LDA is able to capture some notion of the \textit{context} in 
which different usages of the same words occur. As a result, the model is able 
to successfully assign similar topic vectors to quiz questions and lecture 
timepoints specifically when their shared combinations of words denote the same 
deeper meaning.

>>>>

We also examined both models' abilities to match the quiz questions to specific 
temporal intervals within the lectures. The third and fourth rows of 
Supplementary Figure~\ref{fig:compare-bert} display the time series of 
correlations between the embedding weights for each quiz question and each 
timepoint of its corresponding lecture (third row: \textit{Four Fundamental 
Forces}; fourth row: \textit{Birth of Stars}; analogous to main text 
Fig.~\questionCorrs). When computed using LDA embeddings, the questions' 
correlation time series generally appear fairly different from each other (i.e., 
individual questions' traces can be easily visually separated), whereas their 
correlation time series computed using BERT embeddings are all highly similar. 
This indicates that the LDA embeddings are capturing some differences in content 
between the questions about a given lecture that the BERT embeddings are not. 
The time series plots of individual questions' correlations also exhibit clear 
``peaks'' when computed using LDA embeddings, but not when computed using BERT 
embeddings. This indicates that the LDA embeddings are capturing specific 
``matches'' between individual quiz questions and relatively well-defined time 
periods of lecture content, whereas the BERT embeddings appear to blur these 
specific correspondences. Taken together, these comparisons suggest that our LDA 
model trained specifically on the two lectures' contents is sensitive to subtle 
differences between the concepts probed by individual quiz questions, and can 
identify the relevant periods in the lectures where those concepts where 
presented (Fig.~\questionCorrs, 
Supp.~Figs.~\ref{fig:forces-peaks},~\ref{fig:bos-peaks}, 
Supp.~Tab.~\ref{tab:matches}). Meanwhile, BERT, whose embedding weights must be 
able to represent a far broader range of concepts, consequently blurs these 
subtle distinctions between concepts conveyed within the span of a single course 
lecture, thereby affording a ``lower-resolution'' depiction of its contents.

@@

These patterns are reflected in the similarity time series plots shown in panels 
C and D (analogous to main text Fig.~\questionCorrs). The questions' time series 
of topic vector correlations exhibit visually distinct ``peaks'' during specific 
sections of the lectures, outside of which their correlations tend to be 
relatively low. In other words, the set of timepoints whose topic vectors are 
maximally correlated with the topic vector for a given question tend to be not 
only specific, but also temporally contiguous. This indicates that topic vector 
correlation matches questions with sustained, cohesive periods of lecture 
content where the relevant underlying \textit{concepts} are discussed, rather 
than with a sporadic assortment of timepoints where the question and lecture 
content share more superficial similarities. The time series of Jaccard 
similarities, by contrast, are less cleanly structured and do appear to reflect 
such superficial content matches in the more scattered, brief, and weak 
``peaks'' they exhibit. As an example, panel E displays the text of the two quiz 
questions highlighted in green and purple in panels A through D, along with the 
text from the best-matching intervals from the lectures (identified based on
topic vector correlation). Both questions include words that---given the subject 
matter of the two lectures---appear frequently throughout their transcripts 
(e.g., ``\textit{force}'', ``\textit{atom}'', ``\textit{nucleus}'', 
``\textit{cloud}'', ``\textit{dense}'', etc.). While the presence or absence of 
any one (or more) of these words will lead to differences in the degree of 
``match'' based on Jaccard similarity (resulting in the ``jagged'' time series
shown in panel D), LDA is able to capture some notion of the \textit{context} in 
which different usages of the same words occur. As a result, the model is able 
to successfully assign similar topic vectors to quiz questions and lecture 
timepoints specifically when their shared combinations of words denote the same 
deeper meaning.
